---
title: "**Blues in Colors: Police Violence, Racial Representation, and White Attitude Change**"
author: 
  - name: "Chaoyue Wang"
    acknowledgement: "Express your thanks to those who help improve this research along the way."
abstract: "Political behavior has been structured along group identities, and a racial division emerges regarding attitudes toward law enforcement and actions on police brutality. Compared to people of color, white Americans are more supportive of police agencies and more hesitant about reforming policing behavior even in the wake of multiple recent unjustified police-involved homicides. While existing studies attribute such difference to white's unique experiences with law enforcement, excessive white representation in police workforces has received little attention. Linking a nationally representative sample to their local context of racialized police and police violence, this study finds that more representation of black and Hispanic officers greatly enhances the process where white residents reacts to police violence by holding more critical view toward law enforcement. Interestingly, white representation in police has only weak effect of such. Findings here highlights group thinking as a contributing factor to today's racial divide on policing, and implicates how promoting racial diversity in police workforce can facilitate the outset of meaningful conversations on police violence."
format: 
  pdf:
    documentclass: article
    fig_caption: yes
    fontsize: 11pt
    toc: false
    indent: true
    number-sections: true
    fontfamily: libertine
    # mainfont: Crimson
    monofont: inconsolata
    linestretch: 1.5
    geometry: margin = 1in
    include-before:
      - \captionsetup{labelfont = {bf,},font = small} 
    header-includes: 
      - \onehalfspacing 
    fig-cap-location: bottom
    tbl-cap-location: top
    fig-pos: t
    tbl-pos: t
    link-citations: true
    keep-tex: true
    cite-method: biblatex
    biblatexoptions: 
      - style = windycity
      - reflist
      - autocite = inline
      - backend = biber
      - url = false
      - doi = false
      - isbn = false
    colorlinks: true
    citecolor: DodgerBlue4
    linkcolor: DeepSkyBlue4
    urlcolor: SpringGreen4
  html:
    theme: [journal,custom.scss] # lumen journal flatly cosmo
    fig-dpi: 250
    number-sections: true
    code-tools: true
    toc: true
    toc-title: Overview
    toc-location: left
    title-block-banner: '#f0f1f2'
    title-block-banner-color: '#10378c'
    comments:
      hypothesis:
        theme: classic

# csl: chicago.csl
editor: visual
execute:
  echo: false
  warning: false
date: last-modified
date-format: long
editor_options: 
  chunk_output_type: inline
bibliography: references.bib
---

```{r}
#| label: setup
library(here)
library(tidyverse)
library(estimatr)
library(broom)
library(kableExtra)
library(interactions)
library(patchwork)
library(usmap)
library(modelsummary)
library(sjPlot)
library(ggeffects)

load(here('final objects.RData'))
options(modelsummary_format_numeric_latex = "latex")
table.output <- 'gt'
```

# Empirical Strategy

The challenging nature of political behavior writing lies upon not the quantity and complexity demanded for organizing a richly textured yet rigorously ordered narrative, like that in the discipline of historical and cultural analysis, but the rarity, even among the most polished writings in this profession, of a masterful balance of empirical indications and theoretical interpretations where raw, lukewarm data and their statistical derivatives are smoothly situated within a context of vivid, vibrant ideas [@egan2020; @mason2018; @abrajano2015].

Political behavior has been structured along group identities, and a racial division emerges regarding attitudes toward law enforcement and actions on police brutality. Compared to people of color, white Americans are more supportive of police agencies and more hesitant about reforming policing behavior even in the wake of multiple recent unjustified police-involved homicides. While existing studies attribute such difference to white's unique experiences with law enforcement, excessive white representation in police workforce has received little attention. Linking a nationally representative sample to their local context of racialized police and police violence, this study finds that more representation of black and Hispanic officers greatly enhances the process where white residents reacts to police violence by holding more critical view toward law enforcement. Interestingly, white representation in police has only weak effect of such. Findings here highlights group thinking as a contributing factor to today's racial divide on policing, and implicates how promoting racial diversity in police workforce can facilitate the outset of meaningful conversations on police violence.

```{r}
#| label: fig-lemas-density
#| fig-cap: "**Distribution of Racial Presence among Police Departments Surveyed in LEMAS 2016.** On the horizontal axis, a positive value indicates that the corresponding racial group is excessively represented in local police departments, and a negative value the otherwise."
#| fig-width: 11
#| fig-height: 3.65
fig.lemas.density
```

```{r}
#| label: fig-lemas-cover
#| fig-cap: "**Geographic Coverage of LEMAS 2016 at the County Level.** Counties are colored blue where at least one police department within its jurisdiction is surveyed in LEMAS 2016."
#| fig-width: 9.5
#| fig-height: 5
fig.lemas.cover
```

```{r}
#| label: fig-baseline
#| fig-cap: "**the Estiamted Relationship between Racial Imagery of Local Police and White Attitudes on Policing.** Linear predicted values of outcome attitudes are based upon the previous interactional OLS model."
#| fig-width: 10.5
#| fig-height: 3.5

fig.baseline
```

```{r}
#| label: tbl-divides
#| tbl-cap: "Racial Imagery of Local Police Moderates Racial Divides on Police Violence."
modelsummary(
  models.divides,
  # output = 'tables/divide moderation.tex',
  stars = TRUE,output = table.output,
  gof_map = c("nobs", "r.squared"), 
  coef_map = c(
    "white1" = "Racial Divide",
    "white_rep" = 'White Imagery of Police',
    "white_rep:white1" = 'Racial Divide × White Imagery'),
  notes = list('This is the note of your regression table.')) 
```

```{r}
#| label: fig-divides
#| fig-cap: "**Racial Imagery of Local Police Moderates Whites' Attitudinal Response to Police Violence.** Linear predicted values of outcome attitudes are based upon the previous interactional OLS model."
#| fig-width: 11
#| fig-height: 3.65

fig.divides
```

```{r}
#| label: tbl-reaction
#| tbl-cap: "Racial Imagery of Local Police Moderates Racial Reaction to Police Violence."
modelsummary(
  models.reaction,
    stars = TRUE,output = table.output,
    gof_map = c("nobs", "r.squared"),
    coef_map = c( 'white_rep' = 'White Imagery of Police',
                  'pvd' = "Any Police Violence in 2020",
                  'white_rep:pvd' = "Police Violence × White Imagery"),
    notes = list('This is the note of your regression table.') ) 
```

```{r}
#| label: fig-reaction-mod
#| fig-cap: "**Racial Imagery of Local Police Moderates Whites' Attitudinal Response to Police Violence.** Linear predicted values of outcome attitudes are based upon the previous interactional OLS model."
#| fig-width: 11
#| fig-height: 4

fig.reaction.mod
```

```{r}
#| label: fig-racial-component
#| fig-cap: "**Racial Imagery of Local Police Moderates Whites' Attitudinal Response to Police Violence.** Linear predicted values of outcome attitudes are based upon the previous interactional OLS model."
#| fig-width: 11
#| fig-height: 3.65

fig.racial.component
```

```{r}
#| label: tbl-racial.component
#| tbl-cap: "Racial Imagery of Local Police Moderates Whites' Attitudinal Response to Police Violence."
  modelsummary(
  models.racial.component,
  stars = TRUE, output = table.output,
  gof_map = c("nobs", "r.squared"),　
  coef_map = c( 'white_rep' = 'White Imagery of Police',
                'pv_racePV Whites' = "PV Whites",
                'pv_racePV POC' = "PV POC",
                'white_rep:pv_racePV Whites' = "White Imagery × PV Whites",
                'white_rep:pv_racePV POC' = "White Imagery × PV POC"),
  notes = list('This is the note of your regression table.'))
```

# References
